% !TeX program = xelatex
% PCC-NIUC Overleaf-ready manuscript (Simple bibliography version)
\documentclass[11pt]{article}

% Engine-conditional font setup
\usepackage{iftex}
\ifPDFTeX
  \usepackage[utf8]{inputenc}
  \usepackage[T1]{fontenc}
  \usepackage{lmodern}
\else
  \usepackage{fontspec}
  % Use a Unicode font with broad coverage
  \setmainfont{Noto Serif}
\fi

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\usepackage{microtype}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{array}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xcolor}

% Section formatting
\titleformat{\section}{\large\bfseries}{\thesection}{0.6em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}

% Hyperref setup
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
  pdfauthor={PCC-NIUC Team},
  pdftitle={Privacy-Preserving Computing with Non-Interactive Universal Computation}
}

% Title
\title{Privacy-Preserving Computing with Non-Interactive Universal Computation\\\large Cryptographic Enforcement Against Indirect Prompt Injection}
\author{PCC-NIUC Team}
\date{\today}

\begin{document}

\maketitle

% Abstract is extracted from PAPER.md if present; otherwise left empty
\input{abstract.tex}

% Main content converted from PAPER.md
\input{content_simple.tex}

% Simple bibliography section
\section*{References}
\begin{enumerate}
\item OWASP Foundation. OWASP AI Security and Privacy Guide. 2024.
\item NIST. Artificial Intelligence Risk Management Framework (AI RMF 1.0). NIST AI 100-1, January 2023.
\item Necula, G.C. and Lee, P. Safe kernel extensions without run-time checking. ACM SIGOPS Operating Systems Review, 1996.
\item Microsoft. AI Security Risk Assessment Framework. 2024.
\item Meta AI Research. Llama Guard: LLM-based Input-Output Safeguard. 2024.
\item Anthropic. Constitutional AI: Harmlessness from AI Feedback. 2024.
\item Zou, A. et al. Universal and Transferable Adversarial Attacks on Aligned Language Models. 2023.
\item [Additional references - add as needed]
\end{enumerate}

\end{document}


